Concurrent Programming with CXL

### 2020-06-18 book/paper.tex
Latest
Added an exercise

\section*{Exercises}
\begin{enumerate}
\item Figure~\ref{fig:csonebit} presents another correct solution to the
mutual exclusion problem.  It is similar to the one in
Figure~\ref{fig:upflags}, but has a process \emph{back out and try again}
if it finds that the other process is either trying to enter the critical
section or already has.  Compare this algorithm with Peterson's.  What are
the advantages and disadvantages?
\item A colleague of the author asked if the first two assignments in
Peterson's algorithm (setting \texttt{flags[self]})
to \texttt{True} and \texttt{turn} to \texttt{1 - self}) can be reversed.
After all, they are different variables assigned independent values---in a
sequential program one could surely swap the two assignments.
See if you can figure out for yourself if the two assignments can be
reversed.  Then run the program in Figure~\ref{fig:peterson} after reversing
the two assignments and describe in English what happens.
\item Bonus question:
Can you generalize Peterson's algorithm to more than two processes?
\end{enumerate}

@@

CXL can check these invariants as well.  Figure~\ref{fig:tas} already
has the code to check $I_2(i)$.  But how would one go about checking an
invariant like $I_1$.  Invariants must hold for every state.
For $I_2$ we only need an assertion at label \texttt{@cs} because the
premisse is that there is a process at that label.  However, we would
like to check $I_1$ in \emph{every state} (after the variables have
been initialized).

@@

\begin{quote}
Try doing mutual exclusion with other interlock instructions!
\begin{itemize}
\item Find a description for \texttt{compare-and-swap} (CAS)
\index{compare-and-swap}
\index{CAS}
on the internet and
update Figure~\ref{fig:tas}, replacing \texttt{tas}.
\item Other interesting interlock instructions include \texttt{swap} and
\texttt{fetch-and-add}.
\index{fetch-and-add}
\end{itemize}
\end{quote}

@@

In CXL, atomic statements allow you to \emph{implement} your own
low synchronization primitives like test-and-set.  Atomic statements
are not intended to \emph{replace} locks or other synchonization primitives.

@@

@@

Consider a variant of the Dining Philosophers (which were clearly Dutch---who
else need two forks to eat a meal?) that we call Italian Philosophers (who only
need one fork each to eat spaghetti).  Again, there are five philosophers.
This time, however, there is a glass sitting in the middle of the table with only
three forks in it.  A philosopher needs one of those forks to eat, but clearly,
no more than three philosophers can eat at a time.

>>>>

\section*{Exercises}
\begin{enumerate}
\item Figure~\ref{fig:csonebit} presents another correct solution to the
mutual exclusion problem.  It is similar to the one in
Figure~\ref{fig:upflags}, but has a process \emph{back out and try again}
if it finds that the other process is either trying to enter the critical
section or already has.  Compare this algorithm with Peterson's.  What are
the advantages and disadvantages?
\item A colleague of the author asked if the first two assignments in
Peterson's algorithm (setting \texttt{flags[self]})
to \texttt{True} and \texttt{turn} to \texttt{1 - self}) can be reversed.
After all, they are different variables assigned independent values---in a
sequential program one could surely swap the two assignments.
See if you can figure out for yourself if the two assignments can be
reversed.  Then run the program in Figure~\ref{fig:peterson} after reversing
the two assignments and describe in English what happens.
\item Bonus question:
Can you find an inductive invariant for Figure~\ref{fig:csonebit} to prove
it correct?
\item Bonus question:
Can you generalize Peterson's algorithm to more than two processes?
\end{enumerate}

@@

CXL can check these invariants as well.  Figure~\ref{fig:tas} already
has the code to check $I_2(i)$.  But how would one go about checking an
invariant like $I_1$?  Invariants must hold for every state.
For $I_2$ we only need an assertion at label \texttt{@cs} because the
premisse is that there is a process at that label.  However, we would
like to check $I_1$ in \emph{every state} (after the variables have
been initialized).

@@

\section*{Exercises}
\begin{enumerate}
\item Implement an atomic swap operation.  It should take two pointer arguments
and swap the values.
\item Implement a spinlock using the atomic swap operation.
\item Write out the invariants that need to hold and check them using CXL.
\end{enumerate}

@@

In CXL, atomic statements allow you to \emph{implement} your own
synchronization primitives like test-and-set.  Other common examples
include compare-and swap and fetch-and-add.  Atomic statements
are not intended to \emph{replace} locks or other synchonization primitives.
When solving synchronization problems you should not directly use
atomic statement but use the synchronization primitives that are available
to you.  But if you want to design a new synchronization primitive, then
use \texttt{atomic} by all means.

@@

\section*{Exercises}
\begin{enumerate}
\item Write a synchronization module using your own implementation of \texttt{Lock}
(for example, using Peterson's algorithm or using \texttt{swap}), and run the code in
Figure~\ref{fig:incfixed} using your own module.
\end{enumerate}

@@

Consider a variant of the Dining Philosophers (which were clearly Dutch---who
else need two forks to eat a meal?) that we call Italian Philosophers (who only
use one fork each to eat spaghetti---they need the other to speak\footnote{The joke
was provided to me by my dear friend and colleague Lorenzo Alvisi.}).
Again, there are five philosophers.
This time, however, there is a glass sitting in the middle of the table with only
three forks in it.  A philosopher needs one of those forks to eat, but clearly,
no more than three philosophers can eat at a time.
